---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzip the Data and read the csv file
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Finding total number of steps per day
```{r}
agg <- aggregate(steps~date,data=activity, sum, na.rm=TRUE)
colnames(agg)<-c("Date", "NumberOfSteps")
```

Create a histogram of the number of steps per day
```{r}
hist(as.numeric(agg$NumberOfSteps),main = "Histogram of Number of Steps per Day", xlab = "Number of Steps per Day")
```

Finding Mean and Median of the total number of steps
```{r}
mean(as.numeric(agg$NumberOfSteps))
median(as.numeric(agg$NumberOfSteps))
```

## What is the average daily activity pattern?
Make a time series plot
```{r}
interval <- aggregate(steps ~ interval, data=activity, FUN=mean)
plot(interval, type="l", main = "Time Series Plot", xlab = "Interval", ylab="Number of steps")
```

Interval which contains max number of steps:
```{r}
interval$interval[which.max(interval$steps)]
```


## Imputing missing values
Find the total number of missing values (NAs):
```{r}
sum(is.na(activity))
```
Fill missing Data with 5-min interval mean and update the data set
```{r}
for(i in 1:nrow(activity))
{
  if(is.na(activity$steps[i])) 
    activity$steps[i] <- interval$steps[match(activity$interval[i],interval$interval)]
}
```
###Repeat the first part
Finding total number of steps per day
```{r}
agg <- aggregate(steps~date,data=activity, sum, na.rm=TRUE)
colnames(agg)<-c("Date", "NumberOfSteps")
```
Create a histogram of the number of steps per day
```{r}
hist(as.numeric(agg$NumberOfSteps),main = "Histogram of Number of Steps per Day", xlab = "Number of Steps per Day")
```

Finding Mean and Median of the total number of steps
```{r}
mean(as.numeric(agg$NumberOfSteps))
median(as.numeric(agg$NumberOfSteps))
```
**Mean values stays the same but therer is slight difference in meadian value.**


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
activity["day"] <- weekdays(as.Date(activity$date, "%Y-%m-%d"))
activity$day[activity$day %in% c("Saturday", "Sunday")] <-"weekend"
activity$day[activity$day!="weekend"] <-"weekday"
activity$day <- as.factor(activity$day)
head(activity)
```

Make a panel plot containing a time series plot 
```{r}
require("lattice")
plotdata <- aggregate(steps ~ interval + day, data=activity, FUN=mean)
xyplot(plotdata$steps~plotdata$interval|plotdata$day,type='l',layout=c(1,2), xlab = "Interval", ylab = "Steps", main="Time Series")


```

